\Large Taking uncertainty seriously \newline \normalsize A Bayesian approach to word embedding bias estimation

### 2021-08-05 presentationESSLLI21/presentationESSLLI.tex
Adds current presentation version

  pdftitle={Taking uncertainty in word embedding bias estimation seriously ~a Bayesian approach},

@@

@@

\begin{frame}{Cosine-based measures of bias}
\protect\hypertarget{cosine-based-measures-of-bias}{}
\begin{block}{Word embeddings}
\protect\hypertarget{word-embeddings}{}
\begin{itemize}
\item
  Representation of words (or words in contexts) with vectors of real
  numbers
\item
  Built to predict the probability of co-occurence
\end{itemize}

@@

\begin{itemize}
\item
  Geometric interpretation: direction (not length)
\item
  \(\mathsf{cosineDistance}\in (0, 2)\)
\item
  Naive interpretation: proximity corresponds to semantic similarity
  (e.g.~no triangle inequality)
\end{itemize}
\end{block}
\end{frame}

\begin{frame}{Cosine-based measures of bias}
\protect\hypertarget{cosine-based-measures-of-bias-2}{}
\begin{block}{The worry}
\protect\hypertarget{the-worry}{}
In the learning process these models can learn implicit biases that
reflect harmful stereotypical thinking

@@

\begin{frame}{Cosine-based measures of bias}
\protect\hypertarget{cosine-based-measures-of-bias-3}{}
\begin{block}{Stereotypical lists}
\protect\hypertarget{stereotypical-lists}{}
\footnotesize

\begin{itemize}
\item
  feminine occupations: ``homemaker,'' ``nurse,'' ``receptionist,''
  ``librarian,'' etc.
\item
  masculine occupations: ``maestro,'' ``captain,'' ``architect,'' etc.
\end{itemize}

\normalsize
\end{block}

\begin{block}{Visual example}
\protect\hypertarget{visual-example}{}
\vspace{1mm}
\footnotesize

\begin{center}\includegraphics[width=0.6\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-1-1} \end{center}
\normalsize
\end{block}
\end{frame}

\begin{frame}{Cosine-based measures of bias}
\protect\hypertarget{cosine-based-measures-of-bias-4}{}
\begin{block}{Example: direct bias}
\protect\hypertarget{example-direct-bias}{}
\begin{itemize}
\tightlist
\item
  The gender bias of a word \(w\) is its projection on the gender
  direction
  \(\vec{w} \cdot (\overrightarrow{he} - \overrightarrow{she})\)
\end{itemize}

\begin{itemize}
\tightlist
\item
  Given the (ideally) gender neutral words \(N\) and the gender
  direction \(g\) the direct gender bias is:
\end{itemize}

\vspace{-2mm}

\begin{align}
\mathsf{directBias_c(N,g)} & = \frac{\sum_{w\in N}\vert \mathsf{cos}(\vec{w},g)\vert^c}{\vert N \vert }
\end{align}

@@

(\protect\hyperlink{ref-Bolukbasi2016Man}{Bolukbasi, Chang, Zou,
Saligrama, \& Kalai, 2016})
\end{block}
\end{frame}

\begin{frame}{Cosine-based measures of bias}
\protect\hypertarget{cosine-based-measures-of-bias-5}{}
\begin{block}{Example: Word Embedding Association Test (WEAT)}
\protect\hypertarget{example-word-embedding-association-test-weat}{}
\begin{align*}
s(t,A,B) & = \frac{\sum_{a\in A}f(t,a)}{\vert A\vert} - \frac{\sum_{b\in B}f(t,b)}{\vert B\vert}
\\
WEAT(A,B) & = \frac{
\mu\left(\{s(x,A,B)\}_{x\in X}\right) -\mu\left(\{s(y,A,B)\}_{y\in Y}\right) 
}{
\sigma\left(\{s(w,A,B)\}_{w\in X\cup Y}\right)
}
\end{align*}

\begin{itemize}
\item
  \(t\) is a term, \(A, B\) are sets of stereotype attribute words, \$X,
  \(Y\) are protected group words
\item
  For instance, \(X\) might be a set of male names, \(Y\) a set of
  female names, \(A\) might contain stereotypically male-related career
  words, and \(B\) stereotypically female-related family words
\item
  \(s\)-values are used as datapoints in statistical significance tests
\end{itemize}

@@

\begin{frame}{Cosine-based measures of bias}
\protect\hypertarget{cosine-based-measures-of-bias-6}{}
\begin{block}{Our main target: Mean Average Cosine Similarity (MAC)}
\protect\hypertarget{our-main-target-mean-average-cosine-similarity-mac}{}
\begin{align*}
S(t_i, A_j) & = \frac{1}{\vert A_j\vert}\sum_{a\in A_j}\mathsf{cos}(t,a) \\
MAC(T,A) & = \frac{1}{\vert T \vert \,\vert A\vert}\sum_{t_i \in T }\sum_{A_j \in A} S(t_i,A_j)
\end{align*}

@@

\begin{frame}{Cosine-based measures of bias}
\protect\hypertarget{cosine-based-measures-of-bias-7}{}
\begin{block}{Our main target: Mean Average Cosine Similarity (MAC)}
\protect\hypertarget{our-main-target-mean-average-cosine-similarity-mac-1}{}
\footnotesize 
\begin{table}

@@

\begin{frame}{Cosine-based measures of bias}
\protect\hypertarget{cosine-based-measures-of-bias-8}{}
\begin{block}{Known challenges}
\protect\hypertarget{known-challenges}{}
\begin{itemize}
\item
  Gender-direction might be an indicator of bias, but is insufficient.
  After debiasing other non-gendered words can remain in biased
  relations (\protect\hyperlink{ref-Gonen2019lipstick}{Gonen \&
  Goldberg, 2019})
\item
  Methods which involve analogies and their evaluations by human users
  on Mechanical Turk are unreliable
  (\protect\hyperlink{ref-Nissim2020fair}{Nissim, Noord, \& Goot, 2020})
\end{itemize}
\end{block}
\end{frame}

@@

\begin{block}{No design considerations to sample size}
\protect\hypertarget{no-design-considerations-to-sample-size}{}
We investigate the uncertainty that arises from raw sample sizes
\end{block}
\end{frame}

\begin{frame}{Some methodological problems}
\protect\hypertarget{some-methodological-problems-1}{}
\begin{block}{No word class distinction and no control group}
\protect\hypertarget{no-word-class-distinction-and-no-control-group}{}
We make the subclasses clear, add human neutral predicates and neutral
predicates for control

@@

\begin{frame}{The problem with pre-averaging}
\protect\hypertarget{the-problem-with-pre-averaging}{}
\begin{itemize}
\tightlist
\item
  It throws away information about sample sizes
\item
  It removes variation which may result in false confidence
\end{itemize}
\end{frame}

\begin{frame}{Cosine distance and word connection visualization}
\protect\hypertarget{cosine-distance-and-word-connection-visualization}{}
\begin{block}{Analysis of word ``muslim''}
\protect\hypertarget{analysis-of-word-muslim}{}
\begin{center}\includegraphics[width=1.05\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-2-1} \end{center}
\end{block}
\end{frame}

\begin{frame}{Cosine distance and word connection visualization}
\protect\hypertarget{cosine-distance-and-word-connection-visualization-1}{}
\begin{block}{Analysis of word ``priest''}
\protect\hypertarget{analysis-of-word-priest}{}
\begin{center}\includegraphics[width=1.05\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-3-1} \end{center}
\end{block}
\end{frame}

\begin{frame}{Bayesian model}
\protect\hypertarget{bayesian-model}{}
\end{frame}

\begin{frame}{HPDIs}
\protect\hypertarget{hpdis}{}
\begin{block}{Further work}
\protect\hypertarget{further-work}{}
\begin{itemize}
\tightlist
\item
  downstream tasks
\end{itemize}
\end{block}
\end{frame}

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Bolukbasi2016Man}{}%
Bolukbasi, T., Chang, K.-W., Zou, J. Y., Saligrama, V., \& Kalai, A.
(2016). Man is to computer programmer as woman is to homemaker?
Debiasing word embeddings. \emph{CoRR}, \emph{abs/1607.06520}. Retrieved
from \url{http://arxiv.org/abs/1607.06520}

\leavevmode\hypertarget{ref-Caliskan2017semanticsBiases}{}%
Caliskan, A., Bryson, J. J., \& Narayanan, A. (2017). Semantics derived
automatically from language corpora contain human-like biases.
\emph{Science}, \emph{356}(6334), 183--186.
\url{https://doi.org/10.1126/science.aal4230}

@@

>>>>

  pdftitle={Taking uncertainty seriously A Bayesian approach to word embedding bias estimation},

@@

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

@@

\begin{frame}{The plan of the presentation}
\protect\hypertarget{the-plan-of-the-presentation}{}
\begin{itemize}
\tightlist
\item
  Introduction to the topic of bias in word embeddings
\item
  Analysis of methodological problems with MAC implementation
\item
  Explanation of pre-averaging limitations in methods such as WEAT and
  MAC
\item
  Our research and Bayesian approach
\end{itemize}
\end{frame}

\begin{frame}{Cosine-based measures of bias}
\protect\hypertarget{cosine-based-measures-of-bias}{}
\begin{block}{Word embeddings}
\protect\hypertarget{word-embeddings}{}
\begin{itemize}
\item
  Representation of words with vectors of real numbers
\item
  Often built to predict the probability of co-occurence
\end{itemize}

@@

\begin{itemize}
\item
  Geometric interpretation: direction (not length)
\item
  \(\mathsf{cosineDistance}\in (0, 2)\)
\item
  Naive interpretation: proximity corresponds to semantic similarity
\end{itemize}
\end{block}
\end{frame}

\begin{frame}{Cosine-based measures of bias}
\protect\hypertarget{cosine-based-measures-of-bias-2}{}
\begin{block}{The worry}
\protect\hypertarget{the-worry}{}
In the learning process word embeddings can learn implicit biases that
reflect harmful stereotypical thinking

@@

\begin{frame}{Cosine-based measures of bias}
\protect\hypertarget{cosine-based-measures-of-bias-3}{}
\begin{block}{A visual example}
\protect\hypertarget{a-visual-example}{}
\footnotesize

\begin{itemize}
\item
  feminine occupations: ``homemaker,'' ``nurse,'' ``receptionist,''
  ``librarian,'' etc.
\item
  masculine occupations: ``maestro,'' ``captain,'' ``architect,''
  ``boss,'' etc.
\end{itemize}

\normalsize

\vspace{1mm}
\footnotesize

\begin{center}\includegraphics[width=0.6\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-1-1} \end{center}
\normalsize

--\textgreater{}

--\textgreater{}
\end{block}
\end{frame}

\begin{frame}{Cosine-based measures of bias}
\protect\hypertarget{cosine-based-measures-of-bias-4}{}
\begin{block}{Example: Word Embedding Association Test (WEAT)}
\protect\hypertarget{example-word-embedding-association-test-weat}{}
\vspace{1mm}
\footnotesize

\begin{center}\includegraphics[width=0.55\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-2-1} \end{center}
\normalsize

\pause

@@

\begin{itemize}
\item
  \(s_1 = s(he,A,B) = \frac{.6+.7}{2} - \frac{.2+.1}{2} = .65-.15= .5\)
\item
  \(s_2 = s(man,A,B) = .3\),
  \linebreak  \(s_3 = s(woman,A,B) = -.6, s_4 = s(she, A, B) = -.3\)
\end{itemize}

\vspace{-4mm}

\normalsize

\begin{align*}
\mathsf{WEAT}(A,B) & = \frac{\nicefrac{(s_1+s_2)}{2} - \nicefrac{(s_3+s_4)}{2}}{sd(\{s_1,s_2,s_3,s_4\})} \approx 1.93
\end{align*}
\end{block}
\end{frame}

\begin{frame}{Cosine-based measures of bias}
\protect\hypertarget{cosine-based-measures-of-bias-5}{}
\begin{block}{Example: Word Embedding Association Test (WEAT)}
\protect\hypertarget{example-word-embedding-association-test-weat-1}{}
\begin{align*}
s(t,A,B) & = \frac{\sum_{a\in A}f(t,a)}{\vert A\vert} - \frac{\sum_{b\in B}f(t,b)}{\vert B\vert}
\\
WEAT(A,B) & = \frac{
\mu\left(\{s(x,A,B)\}_{x\in X}\right) -\mu\left(\{s(y,A,B)\}_{y\in Y}\right) 
}{
\sigma\left(\{s(w,A,B)\}_{w\in X\cup Y}\right)
}
\end{align*}

\begin{itemize}
\item
  \(t\) is a term, \(A, B\) are sets of stereotype attribute words,
  \(X\), \(Y\) are protected group words
\item
  For instance, \(X\) might be a set of male names, \(Y\) a set of
  female names, \(A\) might contain stereotypically male-related career
  words, and \(B\) stereotypically female-related family words
\item
  \(s\)-values are used as datapoints in statistical significance tests
\end{itemize}

@@

\begin{frame}{Cosine-based measures of bias}
\protect\hypertarget{cosine-based-measures-of-bias-6}{}
\begin{block}{Our main target: Mean Average Cosine Similarity (MAC)}
\protect\hypertarget{our-main-target-mean-average-cosine-similarity-mac}{}
\vspace{1mm}
\footnotesize

\begin{center}\includegraphics[width=0.5\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-3-1} \end{center}
\normalsize

\vspace{-4mm}

\begin{align*}
 s_1 & = s(muslim,A_1) = \frac{cos(muslim,dirty)+cos(muslim,terrorist)}{2}\\
s_2 & = s(muslim,A_2) = \frac{cos(muslim,familiar)+cos(muslim,conservative)}{2}\\
& \vdots \end{align*}

\vspace{-6mm}

\normalsize

\begin{align*}
 MAC(T,A) & = \mathsf{mean}(\{s_i \vert i \in 1, \dots, k\})
\end{align*}
\end{block}
\end{frame}

\begin{frame}{Cosine-based measures of bias}
\protect\hypertarget{cosine-based-measures-of-bias-7}{}
\begin{block}{Our main target: Mean Average Cosine Similarity (MAC)}
\protect\hypertarget{our-main-target-mean-average-cosine-similarity-mac-1}{}
\begin{align*}
S(t_i, A_j) & = \frac{1}{\vert A_j\vert}\sum_{a\in A_j}\mathsf{cos}(t,a) \\
MAC(T,A) & = \frac{1}{\vert T \vert \,\vert A\vert}\sum_{t_i \in T }\sum_{A_j \in A} S(t_i,A_j)
\end{align*}

@@

\begin{frame}{Cosine-based measures of bias}
\protect\hypertarget{cosine-based-measures-of-bias-8}{}
\begin{block}{Our main target: Mean Average Cosine Similarity (MAC)}
\protect\hypertarget{our-main-target-mean-average-cosine-similarity-mac-2}{}
\footnotesize 
\begin{table}

@@

\begin{frame}{Cosine-based measures of bias}
\protect\hypertarget{cosine-based-measures-of-bias-9}{}
\begin{block}{Known challenges}
\protect\hypertarget{known-challenges}{}
\begin{itemize}
\item
  Gender-direction might be an indicator of bias, but is insufficient.
  After debiasing other non-gendered words can remain in biased
  relations (\protect\hyperlink{ref-Gonen2019lipstick}{Gonen \&
  Goldberg, 2019})
\item
  Methods which involve analogies and their evaluations by human users
  on Mechanical Turk are unreliable
  (\protect\hyperlink{ref-Nissim2020fair}{Nissim, Noord, \& Goot, 2020})
\item
  Sensitive to the choice of protected words, capitalization, distance
  measures and embedding methods
  (\protect\hyperlink{ref-zhang2020robustness}{Zhang, Sneyd, \&
  Stevenson, 2020})
\end{itemize}
\end{block}
\end{frame}

@@

\begin{block}{No design considerations to sample size}
\protect\hypertarget{no-design-considerations-to-sample-size}{}
We statistically gauge the uncertainty that arises from raw sample sizes
\end{block}
\end{frame}

\begin{frame}{Some methodological problems}
\protect\hypertarget{some-methodological-problems-1}{}
\begin{block}{No word class distinction and no control group}
\protect\hypertarget{no-word-class-distinction-and-no-control-group}{}
We make the subclasses clear, add human neutral predicates and neutral
predicates for control. We used L2-Reddit corpus and GoogleNews (we
present the results for Reddit for brevity).

@@

\begin{frame}{The problem with pre-averaging}
\protect\hypertarget{the-problem-with-pre-averaging}{}
\begin{block}{Key conceptual issues}
\protect\hypertarget{key-conceptual-issues}{}
\begin{itemize}
\tightlist
\item
  It throws away information about sample sizes
\item
  It ignores variation in the raw data, which leads to false confidence
\end{itemize}
\end{block}

\begin{block}{Our simulation}
\protect\hypertarget{our-simulation}{}
Suppose all similarities for two classes are randomly drawn from the
same distribution, \(\mathsf{Normal}(0,.05)\), you still can get really
high WEAT!
\end{block}
\end{frame}

\begin{frame}{The problem with pre-averaging}
\protect\hypertarget{the-problem-with-pre-averaging-1}{}
\begin{block}{One simulation}
\protect\hypertarget{one-simulation}{}
\vspace{1mm}
\footnotesize

\begin{center}\includegraphics[width=0.7\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-4-1} \end{center}
\normalsize

\vspace{1mm}
\footnotesize

\normalsize
\pause

\vspace{-4mm}

\begin{itemize}
\tightlist
\item
  Raw sd in data is 0.045
\item
  The sd of means is 0.023
\item
  The WEAT score is 1.825
\item
  The largest effect size reported by
  \protect\hyperlink{ref-Caliskan2017semanticsBiases}{Caliskan, Bryson,
  \& Narayanan}
  (\protect\hyperlink{ref-Caliskan2017semanticsBiases}{2017}) is 1.81!
\end{itemize}
\end{block}
\end{frame}

\begin{frame}{The problem with pre-averaging}
\protect\hypertarget{the-problem-with-pre-averaging-2}{}
\begin{block}{50k simulations (same parameters)}
\protect\hypertarget{k-simulations-same-parameters}{}
\vspace{1mm}
\footnotesize

\begin{center}\includegraphics[width=0.8\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-6-1} \end{center}
\normalsize
\end{block}
\end{frame}

\begin{frame}{Cosine distance and connection type}
\protect\hypertarget{cosine-distance-and-connection-type}{}
\begin{block}{Distances for ``muslim''}
\protect\hypertarget{distances-for-muslim}{}
\begin{center}\includegraphics[width=0.8\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-7-1} \end{center}
\end{block}
\end{frame}

\begin{frame}{Cosine distance and connection type}
\protect\hypertarget{cosine-distance-and-connection-type-1}{}
\begin{block}{Distances for ``priest''}
\protect\hypertarget{distances-for-priest}{}
\begin{center}\includegraphics[width=0.8\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-8-1} \end{center}
\end{block}
\end{frame}

\begin{frame}{Advantages of the Bayesian way}
\protect\hypertarget{advantages-of-the-bayesian-way}{}
\begin{itemize}
\tightlist
\item
  Direct impact of sample sizes
\item
  Straightforward interpretation in terms of posterior probabilities
\item
  Freedom to choose granularity level
\item
  More honest risk assessment and decision making
\end{itemize}
\end{frame}

\begin{frame}{Bayesian model}
\protect\hypertarget{bayesian-model}{}
\begin{block}{Choosing priors}
\protect\hypertarget{choosing-priors}{}
\begin{center}\includegraphics[width=1\linewidth]{presentationESSLLI_files/figure-beamer/priorsVis-1} \end{center}
\end{block}
\end{frame}

\begin{frame}[fragile]{Bayesian model architecture}
\protect\hypertarget{bayesian-model-architecture}{}
\vspace{1mm}
\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(rethinking)}
\FunctionTok{options}\NormalTok{(}\AttributeTok{buildtools.check =} \ControlFlowTok{function}\NormalTok{(action) }\ConstantTok{TRUE}\NormalTok{ )}
\NormalTok{religionCoefs }\OtherTok{\textless{}{-}} \FunctionTok{ulam}\NormalTok{(}
  \FunctionTok{alist}\NormalTok{(}
\NormalTok{    cosineDistance }\SpecialCharTok{\textasciitilde{}} \FunctionTok{dnorm}\NormalTok{(mu,sigma),}
\NormalTok{    mu }\OtherTok{\textless{}{-}}\NormalTok{ m }\SpecialCharTok{+}\NormalTok{ co[con],}
\NormalTok{    m }\SpecialCharTok{\textasciitilde{}} \FunctionTok{dnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,.}\DecValTok{5}\NormalTok{),}
\NormalTok{    co[con] }\SpecialCharTok{\textasciitilde{}} \FunctionTok{dnorm}\NormalTok{(}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{),}
\NormalTok{    sigma }\SpecialCharTok{\textasciitilde{}} \FunctionTok{dcauchy}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{)}
\NormalTok{  ),}
  \AttributeTok{data =}\NormalTok{ religion,}
  \AttributeTok{chains=}\DecValTok{2}\NormalTok{ , }\AttributeTok{iter=}\DecValTok{8000}\NormalTok{ , }\AttributeTok{warmup=}\DecValTok{1000}\NormalTok{, }
  \AttributeTok{log\_lik =} \ConstantTok{TRUE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\normalsize
\end{frame}

\begin{frame}{Dataset-level coefficients}
\protect\hypertarget{dataset-level-coefficients}{}
\begin{block}{Religion with 89\%-compatibility intervals (HPDI)}
\protect\hypertarget{religion-with-89-compatibility-intervals-hpdi}{}
\begin{center}
\begin{figure}[!htb]\centering
   \begin{minipage}{0.55\textwidth}
  \includegraphics[width=6cm]{../images/religionCoeffs.jpeg}
   \end{minipage} \begin{minipage}{0.4\textwidth}\footnotesize
   \begin{itemize}
   \item All HPDIs overlap with 0 
   \item Differences between classes are relatively small
   \item Coefficients for Race are similar
   \end{itemize}
   \end{minipage}
  \end{figure}

\end{center}
\end{block}
\end{frame}

\begin{frame}{Dataset-level coefficients}
\protect\hypertarget{dataset-level-coefficients-1}{}
\begin{block}{Gender with 89\%-compatibility intervals (HPDI)}
\protect\hypertarget{gender-with-89-compatibility-intervals-hpdi}{}
\begin{center}
\begin{figure}[!htb]\centering
  \begin{minipage}{0.55\textwidth}
\includegraphics[width=6cm]{../images/genderCoeffs.jpeg}
\end{minipage}
\begin{minipage}{0.4\textwidth}\footnotesize
   \begin{itemize}
   \item Associated and different are away from 0
   \item But they were supposed to be opposites and are very close to each other (co-occurrence?)
   \item Differences between classes are still relatively small
   \end{itemize}
   \end{minipage}
\end{figure}

\end{center}
\end{block}
\end{frame}

\begin{frame}[fragile]{Bayesian model architecture}
\protect\hypertarget{bayesian-model-architecture-1}{}
\vspace{1mm}
\footnotesize

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(rethinking)}
\FunctionTok{options}\NormalTok{(}\AttributeTok{buildtools.check =} \ControlFlowTok{function}\NormalTok{(action) }\ConstantTok{TRUE}\NormalTok{ )}
\NormalTok{religionCoefs }\OtherTok{\textless{}{-}} \FunctionTok{ulam}\NormalTok{(}
  \FunctionTok{alist}\NormalTok{(}
\NormalTok{    cosineDistance }\SpecialCharTok{\textasciitilde{}} \FunctionTok{dnorm}\NormalTok{(mu,sigma),}
\NormalTok{    mu }\OtherTok{\textless{}{-}}\NormalTok{ m[pw] }\SpecialCharTok{+}\NormalTok{ co[con],}
\NormalTok{    m[pw] }\SpecialCharTok{\textasciitilde{}} \FunctionTok{dnorm}\NormalTok{(}\DecValTok{1}\NormalTok{,.}\DecValTok{5}\NormalTok{),}
\NormalTok{    co[con] }\SpecialCharTok{\textasciitilde{}} \FunctionTok{dnorm}\NormalTok{(}\DecValTok{0}\NormalTok{,.}\DecValTok{5}\NormalTok{),}
\NormalTok{    sigma }\SpecialCharTok{\textasciitilde{}} \FunctionTok{dcauchy}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{)}
\NormalTok{  ),}
  \AttributeTok{data =}\NormalTok{ religion,}
  \AttributeTok{chains=}\DecValTok{2}\NormalTok{ , }\AttributeTok{iter=}\DecValTok{8000}\NormalTok{ , }\AttributeTok{warmup=}\DecValTok{1000}\NormalTok{, }
  \AttributeTok{log\_lik =} \ConstantTok{TRUE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\normalsize
\end{frame}

\begin{frame}{Word-level coefficients}
\protect\hypertarget{word-level-coefficients}{}
\begin{figure}[!htb]\centering
  \begin{minipage}{0.55\textwidth}
\includegraphics[width=7.5cm]{../images/visReligionReddit.png}
\end{minipage}
\begin{minipage}{0.4\textwidth}\footnotesize

\vspace{-4cm}

   \begin{itemize}
   \item Most intervals overlap with control groups
   \item Often not too much diffence between associated and different
   \end{itemize}
   \end{minipage}
\end{figure}
\end{frame}

\begin{frame}{Word-level coefficients}
\protect\hypertarget{word-level-coefficients-1}{}
\begin{figure}[!htb]\centering
  \begin{minipage}{0.55\textwidth}
\includegraphics[width=7.5cm]{../images/visGenderReddit.png}
\end{minipage}
\begin{minipage}{0.4\textwidth}\footnotesize

\vspace{-4cm}

   \begin{itemize}
   \item Male attributes strong co-occurrence with female attributes
   \item Sometimes different is stronger than associated
   \item Almost no overlap with control groups
   \end{itemize}
   \end{minipage}
\end{figure}
\end{frame}

\begin{frame}{Word-level coefficients}
\protect\hypertarget{word-level-coefficients-2}{}
\begin{figure}[!htb]\centering
  \begin{minipage}{0.55\textwidth}
\includegraphics[width=7.5cm]{../images/visRaceReddit.png}
\end{minipage}
\begin{minipage}{0.4\textwidth}\footnotesize

\vspace{-4cm}

   \begin{itemize}
   \item A lot of variation between races

   \item Often not much difference between associated and different
   \end{itemize}
   \end{minipage}
\end{figure}
\end{frame}

\begin{frame}{Summary}
\protect\hypertarget{summary}{}
\begin{itemize}
\tightlist
\item
  Introduced to the topic of bias in word embeddings
\item
  Analysed methodological problems with MAC implementation
\item
  Explained of pre-averaging limitations in methods such as WEAT and MAC
\item
  Proposed Bayesian approach
\end{itemize}
\end{frame}

\begin{frame}{Further work}
\protect\hypertarget{further-work}{}
\begin{itemize}
\tightlist
\item
  Including contrasts in Bayesian calculation
\item
  Performance cross-validation in comparison to other methods (regular
  linear regression, KNN, \dots)
\item
  Downstream tasks and connection with intrinsic evaluation
\item
  Testing data from the original Implicit Association Test (IAT)
\item
  Applying uncertainty to WEAT and better word lists
\item
  Looking at other similarity measures
\end{itemize}
\end{frame}

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Caliskan2017semanticsBiases}{}%
Caliskan, A., Bryson, J. J., \& Narayanan, A. (2017). Semantics derived
automatically from language corpora contain human-like biases.
\emph{Science}, \emph{356}(6334), 183--186.
\url{https://doi.org/10.1126/science.aal4230}

@@

\leavevmode\hypertarget{ref-zhang2020robustness}{}%
Zhang, H., Sneyd, A., \& Stevenson, M. (2020). \emph{Robustness and
reliability of gender bias assessment in word embeddings: The role of
base pairs}. Retrieved from \url{http://arxiv.org/abs/2010.02847}
